### 2021-09-13 paper/localpsf.tex
wrote no-breakdown theorem

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = f_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $f_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
\begin{equation}
\label{eq:optimal_v}
v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
\end{equation}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{f} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{f} \in \mathbb{R}^r$ is the vector with $i$th component $f_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation}
\label{eq:S_matrix}
S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
\end{equation}
\end{enumerate}
In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.

@@

>>>>

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = f_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $f_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \theta_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
\begin{equation}
\label{eq:optimal_v}
v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
\end{equation}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{f} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{f} \in \mathbb{R}^r$ is the vector with $i$th component $f_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation}
\label{eq:S_matrix}
S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
\end{equation}
\end{enumerate}
In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, so long as there is no repetition in the points $x_i$. Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $r+1$-dimensional subspace spanned by the functions $\psi_i$, $i=1,\dots,r$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$. In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process solves optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.

@@

\begin{thm}[Procedure for computing $v_*$ does not break down]
	If all the sample points $x_i$, $i=1,\dots,r$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\psi_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\theta_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$. That is, the procedure for computing $v_*$ does not break down. Furthermore, $v_* \in H^2_N(\Omega)$.
\end{thm}

\begin{proof}
	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory. By standard elliptic regularity theory, $\psi_i \in H^1(\Omega)$, so $\psi_i \in L^2(\Omega)$, and so all matrix entries $S_{ij}$ are finite.

	Using the definition of $S$ and basic linear algebra, it is straightforward to show that if $S$ were singular, there would exist constants $a_1, a_2, \dots a_r$, which are not all zero, such that
	\begin{equation*}
		0 = a_1 \psi_1 + a_2 \psi_2 + \dots a_r \psi_r.
	\end{equation*}
	Applying the Laplacian to this equation yields
	\begin{equation*}
		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_r \delta_{x_r} - \left(\sum_{i=1}^r a_i\right)/|\Omega|,
	\end{equation*}
	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function. Thus we must conclude that $S$ is non-singular. By block matrix factorization, non-singularity of $S$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.

	By standard elliptic regularity theory, the functions $\theta_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$. By construction the functions $\theta_i$ have Neumann zero boundary conditions, so $\theta_i \in H^2_N(\Omega)$. Since $v_*$ is a linear combination of the functions $\theta_i$ and the constant function $C(x)=1$, $v_* \in H^2_N(\Omega)$.
\end{proof}
